---
title: "cleaing IV"
author: "Siyang Sun"
date: "April 1, 2016"
output: html_document
---
Clean movie title

```{r}
dfnew<-read.table("/Users/mac/Desktop/EDSP\ /dataclean_3.txt", header =T,sep="\t")
grep("\\(",dfnew$title)
dfnew$title<-as.character(dfnew$title)
dfnew$title[22]<-"Alice in Wonderland"
dfnew$title[131]<-"Cinderella"
dfnew$title[137]<-"Godzilla"
dfnew$title[150]<-"Teenage Mutant Ninja Turtles"
dfnew$title[152]<-"Clash of the Titans"
dfnew$title[189]<-"Les Miserables"
dfnew$title[198]<-"The Intouchables"
dfnew$title[227]<-"Home"
dfnew$title[265]<-"Planet of the Apes"
dfnew$title[285]<-"The Great Gatsby"
dfnew$title[312]<-"Fantastic Four"
dfnew$title[322]<-"A Christmas Carol"
dfnew$title[329]<-"101 Dalmatians"
dfnew$title[414]<-"The Nutty Professor"
dfnew$title[481]<-"Hercules"
dfnew$title[490]<-"RoboCop"
dfnew$title[512]<-"The Day the Earth Stood Still"
dfnew$title[513]<-"The Girl with the Dragon Tattoo"
dfnew$title[589]<-"Fury"
dfnew$title[618]<-"Everest"
dfnew$title[621]<-"Hairspray"
```

Goal 1: Create a label to distinguish super hero movie or not ($hero)
Goal 2: Create a label to distinguish a series moive or not ($series)

```{r}
dfnew$series<-substr(dfnew$title,nchar(dfnew$title),nchar(dfnew$title))
dfnew$series[543]<-0
part<-grep("Part",dfnew$title)
episode<-grep("Episode",dfnew$title)
s23<-union(grep("2",dfnew$series),grep("3",dfnew$series))
s45<-union(grep("4",dfnew$series),grep("5",dfnew$series))
s67<-union(grep("6",dfnew$series),grep("7",dfnew$series))
U<-union(part,episode)
U<-sort(union(union(s23,s45),union(s67,U)),F)
colon<-grep(":",dfnew$title)
se<-union(U,colon)
#134 series movies
dfnew$series[se]<-1
dfnew$series<-ifelse(dfnew$series==1,1,0)
table(dfnew$series)
dfnew$release_date<-NULL
```


```{r}
toma<-read.table("/Users/mac/Desktop/EDSP\ /rt.txt", header =T,sep="\t")
toma$title<-as.character(toma$title)
toma$title[182]<-"Léon: The Professional"
toma$title[228]<-"The City of Lost Children"
toma$title[350]<-"Life Is Beautiful"
toma$title[437]<-"Live Flesh"
toma$title[479]<-"Apt Pupil"
toma$title[617]<-"Aimée & Jaguar"
toma$title[629]<-"Nico and Dani"
toma$title[663]<-"Code Unknown"
toma$title[724]<-"Amélie"
toma$title[734]<-"Kandahar: Le soleil derrière la lune"
toma$title[745]<-"Comment j'ai tué mon père"
toma$title[922]<-"The Best of Youth"
toma$title[957]<-"Noi the Albino"
toma$title1<-toma$title
toma$title1<-gsub("Episode VII","",toma$title1)
toma$title1<-gsub("Episode VI","",toma$title1)
toma$title1<-gsub("Episode V","",toma$title1)
toma$title1<-gsub("Episode III","",toma$title1)
toma$title1<-gsub("Episode I","",toma$title1)
dfnew$title1<-dfnew$title
dfnew$title1<-gsub("Episode III","",dfnew$title1)
dfnew$title1<-gsub("Episode II","",dfnew$title1)
dfnew$title1<-gsub("Episode I","",dfnew$title1)
dfnew$title1<-gsub("-","",dfnew$title1)
dfnew$title1<-gsub(",","",dfnew$title1)
toma$title1<-gsub("-","",toma$title1)
toma$title1<-gsub(",","",toma$title1)
toma$title1<-gsub("\\)","",toma$title1)
dfnew$title1<-tolower(dfnew$title1)
dfnew$title1<-gsub(" ","",dfnew$title1)
toma$title1<-tolower(gsub(" ","",toma$title1))
#match
tomaid<-rep(c(0),nrow(dfnew))
for (i in (1:nrow(dfnew))){
  if (length(grep(dfnew$title1[i],toma$title1))>0){
    tomaid[i]=grep(dfnew$title1[i],toma$title1)  
  }
}
tomaid
tomaid[60]<- 0
tomaid[181]<-0
for (i in (1:nrow(dfnew))){
  if (tomaid[i]>0){
    dfnew$rn[i]<-toma$reviewnum[tomaid[i]]
    dfnew$to_rating[i]<-toma$tomato_rating[tomaid[i]]    
  } 
  else {
    dfnew$rn[i]<-0
    dfnew$to_rating[i]<-0    
  }
}
sum(dfnew$rn>0)
#306/631 has the rotten indexes
```
## superhero movie
1. BV
+ marvel's avengers
+ iron man 
+ guardians of the galaxy
+ the incredibles
+ captain america
+ big hero
+ ant-man
+ unbreakable
+ skyhigh

2. WB
+ the dark knight
+ man of steel
+ batman
+ superman
+ watchmen
+ greenlaten
+ catwoman

3. Sony
+ spiderman
+ hancock

4. Fox
+ deadpool
+ x_men
+ fantastic four

5. Par.
+ ironman
```{r}
b1<-union(grep(("avengers"),dfnew$title1),grep("ironman",dfnew$title1))
b2<-union(grep("guardiansofthegalaxy",dfnew$title1),grep("theincredibles",dfnew$title1))
b3<-union(grep("captainamerica",dfnew$title1),grep("bighero",dfnew$title1))
b4<-union(b3,grep("antman",dfnew$title1))
b5<-union(b1,b2)
bv<-union(b4,b5)
#WB=2
w1<-union(grep("thedarkknight",dfnew$title1),grep("manofsteel",dfnew$title1))
w2<-union(grep("batman",dfnew$title1),grep("superman",dfnew$title1))
wb<-union(w1,w2)

#Sony=3
sony<-union(grep("spiderman",dfnew$title1),grep("hancock",dfnew$title1))
#Fox=4
fox<-union(grep("xmen",dfnew$title1),grep("fantasticfour",dfnew$title1))
#Par.=5
par<-grep("ironman",dfnew$title1)

#create superhero variable
dfnew$sh<-0
dfnew$sh[bv]<-1
dfnew$sh[wb]<-2
dfnew$sh[sony]<-3
dfnew$sh[fox]<-4
dfnew$sh[par]<-5
table(dfnew$sh)
```
Create boxoffice/budget variable :bb 
```{r}
dfnew$bb<-round(dfnew$domestic/dfnew$budget,2)
summary(dfnew$bb)
write.table(dfnew,"/Users/mac/Desktop/EDSP\ /dataclean_4.txt",sep="\t")
```







