---
title: "2017 project data cleaning part (1)"
author: "Siyang Sun"
date: "March 6, 2016"
output: pdf_document
---

1.read raw data from txt file, split data by comma.

```{r}
data =read.delim("/Users/mac/Desktop/EDSP\ /undo_data.txt", header =T,sep="\t")
require(reshape2)
df<-data.frame(matrix(ncol = 20, nrow =663))  
colnames(df)<-c('number','rank','title','studio','worldwide','domestic','overseas',
                'year','release_date','release_year','genre','runtime','mpaa','bugdet',
                'opening_weekend','duration','director','actors','theater','rating')
for (i in 1:633){
  df[i,]<-colsplit(data[i,], ",",c('number','rank','title','studio','worldwide','domestic','overseas',
                               'year','release_date','release_year','genre','runtime','mpaa','bugdet',
                               'opening_weekend','duration','director','actors','theater','rating'))
}
```

2.delete column "year"
3.split genre

```{r}
df$year <- NULL
df$genre <- gsub('Action / Adventure', 'AA', df$genre)
library(stringr)
genrematrix<-str_split_fixed(df$genre, " ", 2)
df$genre1<-genrematrix[,1]
df$genre2<-genrematrix[,2]
df$genre2 <- gsub('/ ', '', df$genre2)
```
4.transfer opening_weekend data to numerical million level,digit=1
##have to check if the ow data is valid or not##
##first delete ow>250 as NA, but still some wrong values here, need to replace where (weeks=NA) to NA.
```{r}
df$opening_weekend<-as.numeric(df$opening_weekend)/1000000
df$ow<-round(df$opening_weekend,digits=1)
df$opening_weekend<- NULL
df$ow[df$ow>=250]<-NA
```
5.transfer date data
5.1 transfer release date
5.2 transfer runtime to mins
5.3 seperate duration
```{r}
#any other way for 5.1?
date<-as.POSIXlt(strptime(df$release_date,"%B %d"))
#format(date,"%m %d")

run <- gsub('min.', '', gsub('hrs. ', ',', df$runtime))
runmatrix<-str_split_fixed(run, ",", 2)
df$runt<-as.numeric(runmatrix[,1])*60+as.numeric(runmatrix[,2])
df$runtime <- NULL
df$duration <- gsub(' / ', ',', df$duration)
duramatrix<-str_split_fixed(df$duration, ",", 2)
df$duration<- NULL
#generate weeks as the number of week (duration)
weeks<-as.numeric(gsub('weeks','',duramatrix[,2]))
df$weeks<-weeks
```
nest steps:
1. fix the ow data
2. clean director, actor data TT
3. consider genre staff keep 2 or 1?
4. consider release date as categorical period variable
5. missing data 
further more:
1. descriptive statistical analysis:
which dependent and independent variable? 
plots
2. regression model
multicollinearity ? 
delete variables?

