---
title: "modeling part I"
author: "Siyang Sun"
date: "April 4, 2016"
output: html_document
---
# Adjusting for ticket price inflation 
Considering the price inflation factor, I adjust movie finanicals (budget,opening weekend, domestic box office) to today's "price" measurement. 
The data I used to make adjustment comes from the website below.
http://www.boxofficemojo.com/about/adjuster.htm
```{r}
inf<-read.csv("/Users/mac/Desktop/EDSP\ /inflation.csv",header =T)
df<-read.table("/Users/mac/Desktop/EDSP\ /dataclean_4.txt", header =T,sep="\t")
df$price<-0
for (i in 2:nrow(inf)){
  df$price[grep(inf$Year[i],df$release_year)]<-inf$Avg..Price[i]    
}
df$price<-gsub("\\$","",df$price)
df$adjusted<-8.7*df$domestic/as.numeric(df$price)
df$adow<-round(8.7*df$ow/as.numeric(df$price),2)
df$adbudget<-round(8.7*df$budget/as.numeric(df$price),2)
df$adjusted<-round(df$adjusted,2)
write.table(df,"/Users/mac/Desktop/EDSP\ /adjusted_data.txt")
df<-read.table("/Users/mac/Desktop/EDSP\ /adjusted_data.txt",header=T)
```
# 1. Linear regression model 
For now, I used a subset of the data to build regression models.
```{r,echo=FALSE}
da<-df
da$number<-NULL
da$worldwide<-NULL
da$overseas<-NULL
da$title1<-NULL
da$rank<-NULL
da$title<-NULL
da$genre1<-NULL
da$genre2<-NULL  
da$director1<-NULL
da$director2<-NULL
da$star1<-NULL
da$star2<-NULL
da$bb<-NULL
da$domestic<-NULL
da$price<-NULL
da$adjusted<-df$adjusted
da$adow<-NULL
da$adbudget<-NULL
da$rn[which(da$rn==0)]<-NA
da$to_rating[which(da$to_rating==0)]<-NA
write.table(da,"/Users/mac/Desktop/EDSP\ /subset_res.txt")
```
## Outcome variable: domestic box office (adjusted)
### 1. treating studio genre mpaa per as categorical variables
```{r}
#baseline model
res0<-lm(adjusted~budget,na.omit(da))
summary(res0)
##how to deal with missing data here?
#stepwise regression
da$release_year<-factor(da$release_year)
res1<-lm(da$adjusted~.,da)
summary(res1)
```
This model gives a R^2=78.6%.
However, does year ever makes sense when we would like to build a precitive model?
Try deleting release_year variable here. 
```{r}
da$release_year<-NULL
res2<-lm(adjusted~.,da)
summary(res2)
## This time the R square gets a little bit lower, which is 73.6%.
#omitted missing data
dn<-na.omit(da)
#That reduces my dataset by the size of 120 observations.
new1<-lm(adjusted~.,dn)
summary(new1)
## Is there anyway to do not omit data and conduct the stepwise procedure? 
library(MASS)
step <- stepAIC(new1, direction="backward")
step$anova
new1step<-lm(adjusted ~ studio + genre + mpaa + ow + theater+weeks+rn+sh,dn)
summary(new1step)
par(mfrow=c(1,1))
plot(new1step,cex=0.5)
abline(new1step)
#A little bit improvement. The R square is 75.7% now. 
```
Final model:
new1step<-lm(adjusted ~ studio + genre + mpaa + ow + theater+weeks+rn+sh,dn)
Adjusted R-squared:  0.757 
### 2. exclude the superhero movie?
```{r}
res3<-lm(adjusted~.,da[da$sh==0,])
summary(res3)
```
A little bit improvement, R^2 75.9%.
### 3. delete high correlated varaibles to increase sample sizes and avoid collinearity.
Try delete theater,opening weekend and weeks (too many NAs inside, highly correlated)
```{r}
vars<-names(da) %in% c("theater","ow","weeks")
dv<-da[!vars]
##??? why it does not work?
#res4<-lm(adjusted ~budget,dv)
#summary(res4)
```
##2. Outcome variable: openging weekend 
```{r}
res5<-lm(ow~.-adjusted-ow,na.omit(da))
summary(res5)
step <- stepAIC(res5, direction="backward")
step$anova
res5step<-lm(ow ~ studio + mpaa + theater + weeks + per + budget + rn + sh,da)
summary(res5step)
```
R^2 is 60.5%, not better.

#2.Regression decision Tree
```{r}

```

